Supplementary Information for: Crosslinguistic Word Orders Enable an Efficient Tradeoff between Memory and Surprisal

### 2020-05-10 writeup/control-ngrams.tex
code

\subsection{Method}
We use a version of Kneser-Ney Smoothing.
For a sequence $w_1\dots w_k$, let $N(w_{1\dots k})$ be the number of times $w_{1\dots k}$ occurs in the training set.
The unigram probabilities are estimated as
\begin{equation}
	p_1(w_t) :=   \frac{N(w_t) + \delta}{|Train| + |V| \cdot \delta}
\end{equation}
where $\delta \in \mathbb{R}_+$ is a hyperparameter.
Here $|Train|$ is the number of tokens in the training set, $|V|$ is the number of types occurring in train or held-out data.
Higher-order probabilities $p_t(w_t|w_{0 \dots t-1})$ are estimated recursively as follows.
Let $\gamma > 0$ be a hyperparameter.
If $N(w_{0 \dots t-1}) < \gamma$, set
\begin{equation}
	p_t(w_t|w_{0 \dots t-1}) := p_{t-1}(w_t|w_{1\dots t-1})
\end{equation}
Otherwise, we interpolate between $t$-th order and lower-order estimates:
\begin{equation}
	p_t(w_t|w_{0 \dots t-1}) :=  \frac{\operatorname{max}(N(w_{0\dots t}) - \alpha, 0.0) + \alpha \cdot \#\{w : N(w_{0 \dots t-1}w) > 0\} \cdot p_{t-1}(w_t|w_{1\dots t-1})}{N(w_{0\dots t-1})}
\end{equation}
where $\alpha \in [0,1]$ is also a hyperparameter.
(CITE) show that this definition results in a well-defined probability distribution, i.e., $\sum_{w \in V} p_t(w|w_{0 \dots t-1}) = 1$.

@@

Hyperparameters $\alpha, \gamma, \delta$ are tuned with the same strategy as for the neural network models.

\subsection{Results}

\begin{center}
\begin{longtable}{ccccccccccccccclll}
\input{tables/medians_ngrams_0.tex}
\input{tables/medians_ngrams_1.tex}
\input{tables/medians_ngrams_2.tex}
\input{tables/medians_ngrams_3.tex}
\end{longtable}
	\captionof{figure}{Medians (estimated using n-gram models): For each memory budget, we provide the median surprisal for real and random languages. Solid lines indicate sample medians for ngrams, dashed lines indicate 95 \% confidence intervals for the population median. Green: Random baselines; blue: real language; red: maximum-likelihood grammars fit to real orderings.}\label{tab:medians_ngrams}
\end{center}

>>>>

Here we show that the results of Study 2 remain robust when estimating surprisal with a simple n-gram model instead of recurrent neural networks.

\subsubsection{Method}
We use a version of Kneser-Ney Smoothing \citep{kneser-improved-1995}.
For a sequence $w_1\dots w_k$, let $N(w_{1\dots k})$ be the number of times $w_{1\dots k}$ occurs in the training set.
The unigram probabilities are estimated as
\begin{equation}
	p_1(w_t) :=   \frac{N(w_t) + \delta}{|Train| + |V| \cdot \delta}
\end{equation}
where $\delta \in \mathbb{R}_+$ is a hyperparameter.
Here $|Train|$ is the number of tokens in the training set, $|V|$ is the number of types occurring in train or held-out data.
Higher-order probabilities $p_t(w_t|w_{0 \dots t-1})$ are estimated recursively as follows.
Let $\gamma > 0$ be a hyperparameter.
If $N(w_{0 \dots t-1}) < \gamma$, set
\begin{equation}
	p_t(w_t|w_{0 \dots t-1}) := p_{t-1}(w_t|w_{1\dots t-1})
\end{equation}
Otherwise, we interpolate between $t$-th order and lower-order estimates:
\begin{equation}
	p_t(w_t|w_{0 \dots t-1}) :=  \frac{\operatorname{max}(N(w_{0\dots t}) - \alpha, 0.0) + \alpha \cdot \#\{w : N(w_{0 \dots t-1}w) > 0\} \cdot p_{t-1}(w_t|w_{1\dots t-1})}{N(w_{0\dots t-1})}
\end{equation}
where $\alpha \in [0,1]$ is also a hyperparameter.
\citet{kneser-improved-1995} show that this definition results in a well-defined probability distribution, i.e., $\sum_{w \in V} p_t(w|w_{0 \dots t-1}) = 1$.

@@

Hyperparameters $\alpha, \gamma, \delta$ are tuned using the held-out set, with the same strategy as for the neural network models.

\subsubsection{Results}

\begin{center}
\begin{longtable}{ccccccccccccccclll}
\input{tables/medians_ngrams_0.tex}
\input{tables/medians_ngrams_1.tex}
\input{tables/medians_ngrams_2.tex}
\input{tables/medians_ngrams_3.tex}
\end{longtable}
	\captionof{figure}{Memory-surprisal tradeoff curves (estimated using n-gram models): For each memory budget, we provide the median surprisal for real and random languages. Solid lines indicate sample medians for ngrams, dashed lines indicate 95 \% confidence intervals for the population median. Green: Random baselines; blue: real language; red: maximum-likelihood grammars fit to real orderings.}\label{tab:medians_ngrams}
\end{center}
